---
title: "GPSM Simulations"
author: "Troy Rowan"
date: "9/19/2019"
output: html_document
---

```{r setup, include=FALSE}
library(cowplot)
library(reshape2)
library(qvalue)
library(dplyr)
library(tidyr)
library(readr)
library(purrr)
library(forcats)
library(stringr)
knitr::opts_chunk$set(echo = TRUE)
source("plotting_functions.R")
```

##Tracking AF over time

```{r}
scenario100_qtl_gv = read_csv("gpsm_runs/scenario77/scenario77.gv_pop.qtl_trajectories.csv") %>% 
  left_join(., readgwas("gpsm_runs/scenario77/scenario77.gv_pop.gpsm.assoc.txt"))
scenario100_qtl_pheno = read_csv("gpsm_runs/scenario77/scenario77.pheno_pop.qtl_trajectories.csv") %>% 
  left_join(., readgwas("gpsm_runs/scenario77/scenario77.pheno_pop.gpsm.assoc.txt"))
scenario100_qtl_rand = read_csv("gpsm_runs/scenario77/scenario77.rand_pop.qtl_trajectories.csv") %>% 
  left_join(., readgwas("gpsm_runs/scenario77/scenario77.rand_pop.gpsm.assoc.txt"))

scenario100_qtl_gv %>%
  top_n(., 20, abs(effect)) %>% 
  melt(id = c("chr", "pos", "effect", "rs"), value.name = "AF") %>% 
  mutate(generation = as.numeric(str_remove(variable, "^X"))) %>% 
  ggplot(aes(x = generation, y = AF, color = rs))+
  geom_smooth(se = FALSE)+
  scale_color_viridis_d()+
  labs(x = "Generation", title = "True BV Selection Top QTL")

scenario100_qtl_pheno %>%
  top_n(., 20, abs(effect)) %>% 
  melt(id = c("chr", "pos", "effect", "rs"), value.name = "AF") %>% 
  mutate(generation = as.numeric(str_remove(variable, "^X"))) %>% 
  ggplot(aes(x = generation, y = AF, color = rs))+
  geom_smooth(se = FALSE)+
  scale_color_viridis_d()+
  labs(x = "Generation", title = "Phenotypic Selection Top QTL")

scenario100_qtl_rand %>%
  top_n(., 20, abs(effect)) %>% 
  melt(id = c("chr", "pos", "effect", "rs"), value.name = "AF") %>% 
  mutate(generation = as.numeric(str_remove(variable, "^X"))) %>% 
  ggplot(aes(x = generation, y = AF, color = rs))+
  geom_smooth(se = FALSE)+
  scale_color_viridis_d()+
  labs(x = "Generation", title = "Random Selection Top QTL")
#############################################################
scenario100_qtl_gv %>%
  filter(-log10(q) > 0.8) %>% 
  top_n(., 20, abs(-log10(q))) %>%
  melt(id = c("chr", "pos", "effect", "rs", "af", "p_score", "q"), value.name = "AF") %>% 
  mutate(generation = as.numeric(str_remove(variable, "^X"))) %>% 
  ggplot(aes(x = generation, y = AF, group = rs, color = -log10(q)))+
  geom_smooth(se = FALSE)+
  scale_color_viridis_c()+
  labs(x = "Generation", title = "True BV Selection Top GPSM Hits")

scenario100_qtl_pheno %>%
  filter(-log10(q) > 0.8) %>%
  top_n(., 20, abs(-log10(q))) %>%
  melt(id = c("chr", "pos", "effect", "rs", "af", "p_score", "q"), value.name = "AF") %>% 
  mutate(generation = as.numeric(str_remove(variable, "^X"))) %>% 
  ggplot(aes(x = generation, y = AF, group = rs, color = -log10(q)))+
  geom_smooth(se = FALSE)+
  scale_color_viridis_c()+
  labs(x = "Generation", title = "Phenotypic Selection Top GPSM Hits")

scenario100_qtl_rand %>%
  filter(-log10(q) > 0.8) %>% 
  top_n(., 20, abs(-log10(q))) %>%
  melt(id = c("chr", "pos", "effect", "rs", "af", "p_score", "q"), value.name = "AF") %>% 
  mutate(generation = as.numeric(str_remove(variable, "^X"))) %>% 
  ggplot(aes(x = generation, y = AF, group = rs, color = -log10(q)))+
  geom_smooth(se = FALSE)+
  scale_color_viridis_c()+
  labs(x = "Generation", title = "Random Selection Top GPSM Hits")


```

```{r}
readgwas("gpsm_runs/scenario86/scenario86.gv_pop.gpsm.assoc.txt") %>% 
  ggmanhattan(value = q)

readgwas("gpsm_runs/scenario86/scenario86.pheno_pop.gpsm.assoc.txt") %>% 
  ggmanhattan(value = q)
```


```{r}
var_explained = read_tsv("gpsm_runs/scenario77/scenario77.gv_pop.gpsm.assoc.txt") %>% 
  mutate(cum_pos = (chr-1)*1000+ps) %>% 
  mutate(rs = paste(chr, ps, sep = ":")) %>% 
  left_join(., select(read_csv("gpsm_runs/scenario77/scenario77.gv_pop.qtl_trajectories.csv"), rs, effect), by = "rs")


var_explained %>% 
  filter(is.na(effect)) %>% 
  ggplot(aes(x = cum_pos, y = beta))+
    geom_point(shape = 1, alpha = 0.5)+
    geom_point(data = filter(var_explained, !is.na(effect)), aes(x = cum_pos, y = beta, color = effect))+
    scale_color_viridis_c()
```

#Identifying the average magnitude of allele frequency change in a population not under selection
* Here I'm using the randomly-mated populations to quantify this
* Use the completely polygenic architecture (ie 10,000 causal SNPs) - Scenario 9
* Since we keep track of this information, we can calculate changes 
* This will tell us what we should expect in an effective population of around 180 or so over the course of 20 generations
    + Or maybe we should just be looking for the distribution in a couple of
* We see MUCH larger changes than what we'd expect under the "drift variance "
```{r}
run9 = read_csv("gpsm_runs/scenario9/scenario9.rand_pop.qtl_trajectories.csv") %>% 
  mutate(change1 = X2 - X1,
         change2 = X3 - X2,
         change3 = X4 - X3,
         change4 = X5 - X4,
         change5 = X6 - X5,
         change6 = X7 - X6,
         change7 = X8 - X7,
         change8 = X9 - X8,
         change9 = X10 - X9,
         change10 = abs(X11 - X10)) %>% 
  select(rs, contains("change")) %>% 
  melt(id = "rs")

read_csv("gpsm_runs/scenario9/scenario9.rand_pop.qtl_trajectories.csv") %>% 
  mutate(change1 = X2 - X1,
         dvar = (X1*(1-X1))/(1.5*180)) %>% 
  select(rs, change1, dvar) %>% View()

ggplot(run9, aes(value))+
  geom_histogram()

```

